The problem and promise of scale in multilayer animal social networks

### 2020-02-15 paper/manuscript.Rmd
fin caps

### Temporal scale
#### Number of observations
Studies of social network analysis vary in the number and frequency of
observations as well as the data collection technique used to generate networks
[@Davis_2018; @Webber_2019]. For example, GPS data is commonly collected at a
fixed rate, e.g. every $x$ minutes or hours, continuously throughout the study
period. Fix rate is a reflection of number of observations an individual would
have been observed and recorded in traditional ethological studies.We
investigated the influence of the number of observations on resulting multilayer
networks. We randomly selected a maximum 500 timesteps for each season to
represent all possible observations. We regenerated multilayer networks and
calculated associations of individuals with 25 additional observations at each
iteration. Subsequent iterations included the previous set of observations to
mimic collection of observational data. <!-- TODO: bah conclude / this sentence
falls off a cliff -->

@@

<!-- Figure: metrics by landcover resolution -->
(ref:lcres) Varying spatial scale in landcover resolution. For each landcover resolution, multilayer networks were constructed of caribou social associations (*Rangifer tarandus*, n = `r DT[, uniqueN(get(idcol))]` females) on Fogo Island, Newfoundland in summer 2017 and winter 2018 across three habitat classes (forage, forest, and open). A) Mean graph strength of individuals in winter forest and open habitats increased consistently while winter forage varied with a peak at 250 m and valley at 500 m. Individuals in summer layers had consistent weak connections. B) Mean layer relevance showed sparse connectivity in summer compared to winter networks and a decrease in winter forage layer relevance at higher landcover resolutions. C) Similarity of landcover network layers between seasons (e.g. open summer compared to open winter) was relatively stable for forest and open, except for a sharp decrease in open at 100 m and 250 m. Note that layer similarity in summer forage is not shown because individuals did not interact at landcover resolutions greater than 30 m and therefore layer similarity could not be calculated. D) Mean connective redundancy across individuals showed a slight decreasing trend with increasing landcover resolution.

@@

<!-- Figure: metrics by number of observations -->
(ref:nobs) Varying scale in number of observations. For each iteration, a number of observations (10-485 observations in steps of 25) were selected from the complete data set to construct multilayer networks of caribou social associations (*Rangifer tarandus*, n = `r DT[, uniqueN(get(idcol))]` females) on Fogo Island, Newfoundland in summer 2017 and winter 2018 across three habitat classes (forage, forest, and open). A) Mean graph strength across individuals varied for winter layers becoming relatively stable at 100 observations and remained weak throughout for summer layers. B) Mean layer relevance across individuals increased sharply for winter layers between ~10 m and ~50 m observations before stabilizing. Summer layers had consistent low layer relevance. C) Similarity of network layers across habitat classes and between seasons (e.g. open summer compared to open winter) was relatively consistent across number of observations for forest and open habitats. Note, forage habitats are not shown given no individuals interacted in the summer forage layer. D) Mean connective redundancy across individuals increased rapidly between ~10 m and ~50 m observations and then remained constant.

@@

<!-- Figure: Window length -->
(ref:winlen) Varying temporal scale in time window length. For each time window length (40 to 100 days in length increasing by 5 days in sequence), multilayer networks were constructed of caribou social associations (*Rangifer tarandus*, n = `r DT[, uniqueN(get(idcol))]` females) on Fogo Island, Newfoundland in summer 2017 and winter 2018 across three habitat classes (forage, forest, and open). A) Mean graph strength across individuals was relatively stable over increasing time window lengths for winter layers, though there was more variation in strength between landcover types as window length increased. Summer layers an increase in strength for open and forage habitats. B) Mean layer relevance was stable for winter layers and increased for summer layers across habitat classes. C) Similarity of layers across habitat classes and between seasons (e.g. open summer compared to open winter) was relatively stable. D) Mean connective redundancy across individuals increased marginally with increasing window length.

@@

<!-- Figure: Window position -->
(ref:winpos) Varying temporal scale in time window position. For each time window position (shifting the start date of the time window used to define seasons 1 to 48 days from the original), multilayer networks were constructed of caribou social associations (*Rangifer tarandus*, n = `r DT[, uniqueN(get(idcol))]` females) on Fogo Island, Newfoundland in summer 2017 and winter 2018 across three habitat classes (forage, forest, and open). A) Mean graph strength across individuals increased in summer network layers at window position 25 and increased between window position 20 and 40 for winter network layers. B) Mean layer relevance slightly decreased for winter layers and increased for summer layers. C) Similarity of layers across habitat classes and between seasons (e.g. open summer compared to open winter) was relatively stable until window position ~28 where forage and open layer similarity decreased. D) Mean connective redundancy across individuals increased marginally with advancing time window position.

@@

The integration of behavioural and ecological theory has shown the value of
considering the importance of various ecological disciplines to better
understand and predict animal social networks, including demography
[@Shizuka_2019], evolutionary and quantitative genetics [@Fisher_2017], disease
ecology [@Silk_2017], and animal personality [@Sih_2018]. While the number of
studies integrating social network theory into a broader array of behavioural
and ecological sub-disciplines is impressive, only recently has the role of
animal space use been considered in describing the causes and consequences of
the social environment in the context of (1) phenotypes and individual fitness
[@Webber_2018], (2) movement ecology and collective movement [@Bode_2011], and
(3) habitat selection and space use [@He_2019]. Importantly, much of the work
linking space use and the social environment has been theoretical and empirical
studies are lacking [but see @Peignier_2019; @Strandburg_Peshkin_2015; @Strandburg_Peshkin_2017]. Here, we discuss the role
mutlilayer networks<!-- TODO: withinwithin--> in the emerging integration of
social and spatial ecology within the behavioural ecology literature.

>>>>

### Temporal scale
#### Number of observations
Studies of social network analysis vary in the number and frequency of
observations as well as the data collection technique used to generate networks
[@Davis_2018; @Webber_2019]. For example, GPS data is commonly collected at a
fixed rate, e.g. every $x$ minutes or hours, continuously throughout the study
period. Fix rate is a reflection of number of observations an individual would
have been observed and recorded in traditional ethological studies.We
investigated the influence of the number of observations on resulting multilayer
networks. We randomly selected a maximum 485 timesteps for each season to
represent all possible observations. We regenerated multilayer networks and
calculated associations of individuals with 25 additional observations at each
iteration (10 - 485 observations). Subsequent iterations included the previous set of observations to
mimic collection of observational data. <!-- TODO: bah conclude / this sentence
falls off a cliff -->

@@

<!-- Figure: metrics by landcover resolution -->
(ref:lcres) Varying spatial scale in landcover resolution. For each landcover resolution, multilayer networks were constructed of caribou social associations (*Rangifer tarandus*, n = `r DT[, uniqueN(get(idcol))]` females) on Fogo Island, Newfoundland in summer 2017 and winter 2018 across three habitat classes (forage, forest, and open). A) Mean graph strength of individuals in summer layers had consistent weak connections. Individuals in winter forest and open habitats increased consistently while winter forage varied with a peak at 250 m and valley at 500 m. B) Mean layer relevance showed sparse connectivity in summer compared to winter networks and a decrease in winter forage layer relevance at higher landcover resolutions. C) Similarity of landcover network layers between seasons (e.g. open summer compared to open winter) orage and forest habitats but showed more variation in open habitats with increasing social thresholds. Note that layer similarity in forage is not shown because individuals did not interact at landcover resolutions greater than 30 m during the summer in forage habitats, so layer similarity could not be calculated. D) Mean connective redundancy across individuals showed a slight decreasing trend with increasing landcover resolution.

@@

<!-- Figure: metrics by number of observations -->
(ref:nobs) Varying scale in number of observations. For each iteration, a number of observations (10-485 observations in steps of 25) were selected from the complete data set to construct multilayer networks of caribou social associations (*Rangifer tarandus*, n = `r DT[, uniqueN(get(idcol))]` females) on Fogo Island, Newfoundland in summer 2017 and winter 2018 across three habitat classes (forage, forest, and open). A) Mean graph strength across individuals remained weak throughout for summer layers and varied for winter layers becoming relatively stable by 100 observations. B) Mean layer relevance across individuals in summer layers had consistent low layer relevance but increased sharply for winter layers between ~10 m and ~50 m observations before stabilizing. C) Similarity of network layers across habitat classes and between seasons (e.g. open summer compared to open winter) was relatively consistent across number of observations for forest and open habitats. Note, forage habitats are not shown given no individuals interacted in the summer forage layer. D) Mean connective redundancy across individuals increased rapidly between ~10 m and ~50 m observations and then remained constant.

@@

<!-- Figure: Window length -->
(ref:winlen) Varying temporal scale in time window length. For each time window length (40 to 100 days in length increasing by 5 days in sequence), multilayer networks were constructed of caribou social associations (*Rangifer tarandus*, n = `r DT[, uniqueN(get(idcol))]` females) on Fogo Island, Newfoundland in summer 2017 and winter 2018 across three habitat classes (forage, forest, and open). A) Mean graph strength across individuals in summer layers was weak but stable across increasing lengths of time except for a slight increase in strength for open and forage habitats when the season lasted more than 70 days. Mean graph strength was relatively stable over increasing time window lengths for winter layers, though there was more variation in strength between landcover types as window length increased. B) Mean layer relevance increased for summer layers across habitat classes and was stable for winter layers. C) Similarity of layers across habitat classes and between seasons (e.g. open summer compared to open winter) was variable with increasing window lengths, particularly for forage and open habitat types. D) Mean connective redundancy across individuals increased marginally with increasing window length.

@@

<!-- Figure: Window position -->
(ref:winpos) Varying temporal scale in time window position. For each time window position (shifting the start date of the time window used to define seasons 1 to 48 days from the original), multilayer networks were constructed of caribou social associations (*Rangifer tarandus*, n = `r DT[, uniqueN(get(idcol))]` females) on Fogo Island, Newfoundland in summer 2017 and winter 2018 across three habitat classes (forage, forest, and open). A) Mean graph strength across individuals increased in summer network layers at window position 25 and increased between window position 20 and 40 for winter network layers. B) Mean layer relevance increased for summer layers and slightly decreased for winter layers. C) Similarity of layers across habitat classes and between seasons (e.g. open summer compared to open winter) was relatively stable until window position ~28 where forage and open layer similarity decreased. D) Mean connective redundancy across individuals increased marginally with advancing time window position.

@@

The integration of behavioural and ecological theory has shown the value of
considering the importance of various ecological disciplines to better
understand and predict animal social networks, including demography
[@Shizuka_2019], evolutionary and quantitative genetics [@Fisher_2017], disease
ecology [@Silk_2017], and animal personality [@Sih_2018]. While the number of
studies integrating social network theory into a broader array of behavioural
and ecological sub-disciplines is impressive, only recently has the role of
animal space use been considered in describing the causes and consequences of
the social environment in the context of (1) phenotypes and individual fitness
[@Webber_2018], (2) movement ecology and collective movement [@Bode_2011], and
(3) habitat selection and space use [@He_2019]. Importantly, much of the work
linking space use and the social environment has been theoretical and empirical
studies are lacking [but see @Peignier_2019; @Strandburg_Peshkin_2015; @Strandburg_Peshkin_2017]. Here, we discuss the role of scale and 
mutlilayer networks in the emerging integration of
social and spatial ecology within the behavioural ecology literature.
